### 2023-01-05 README.rmd
updated readme

---
output: github_document
---

<!-- README.md is generated from README.Rmd. -->

# contratrain-muscle-paper

## Authors

Håvard Hamarsland
<a href="https://orcid.org/0000-0002-7343-2774">
<img alt="ORCID logo" src="https://info.orcid.org/wp-content/uploads/2019/11/orcid_16x16.png" width="16" height="16" /></a>,
Kristian Lian,
Chidimma Phoebe Echebiri,
Mads-Henrik Hafsmo,
Tomas Varpestuen,
Stein-Olaf Olsen,
Chris Sylstad,
Max Ullrich,
Stian Ellefsen
<a href="https://orcid.org/0000-0003-0467-2018">
<img alt="ORCID logo" src="https://info.orcid.org/wp-content/uploads/2019/11/orcid_16x16.png" width="16" height="16" /></a>,
Daniel Hammarström
<a href="https://orcid.org/0000-0001-8360-2100">
<img alt="ORCID logo" src="https://info.orcid.org/wp-content/uploads/2019/11/orcid_16x16.png" width="16" height="16" /></a>

## Repository organization

`/resources`

- bibliography.bib
- csl-files

`/data`

- contains raw data not included in the data package

`/data-gen/`

- contains generated data stored for further analyses

`contratrain-manuscript.qmd`

- The manuscript source file

`/figures`

- Figures source files and output