### 2023-02-01 draft-styles.RMD
updated change analyses and figures

---
title: "Untitled"
author: "Kristian Lian"
date: "31 1 2023"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r packages}
library(tidyverse);library(readxl);library(lme4);library(emmeans);library(dplyr);library(cowplot);library(knitr);library(broom);library(rmarkdown);library(tinytex);library(tidyr);library(arsenal);library(magrittr);library(dabestr)
```

# Glucose ingestion before and after resistance training does not augment ribosome biogenesis in young moderately trained adults

## Disadvantages due to Covid-19

Due to restrictions regarding social distancing and lock downs caused by the Covid-19 pandemic,
we were not able to recruit and include as many participants as we planned to the study. The
pandemic was also a contributing factor to the endocrine analysis not being completed, as they
were supposed to be performed at Sykehuset Innlandet Hospital Trust.

## Methods

This data set is a part of a larger project involving several investigators and other outcomes not
covered here. All participants were informed about the potential discomforts and risks associated
with the study and gave their informed consent prior to study enrollment. The project was
approved by the regional ethical committee (REK, ID nr. 153628), pre registered at
clinicaltrials.gov (Identifier: NCT04545190) and conducted according to the Helsinki declaration

### Participants

Sixteen male and female participants (20-33yrs, Tab 1) were recruited to the study through
facebook advertisement and word of mouth and taken through the selection process (Fig 1). The
eligibility criteria were non-smokers and moderately trained (i.e. 2-8 resistance training sessions
per 14 days for the last six months). Exclusion criteria were previous injury leading to impaired
strength, inability to perform resistance training and symptoms, and a medical record of
metabolic disorders including hyperglycemia, i.e. fasting venous plasma glucose ≥6.1 mmol/L
and/or 2-hour glucose tolerance ≥7.8 mmol/L, and/or HbA1c >42 mmol/mol. Our goal was to
recruit 20 participants to the study, however due to the advents of Covid-19, we were not able to
do so. Sixteen participants commenced the intervention, during which three dropped out. One
participant had a sick child, and was unable to resume the intervention, two participants
experienced muscular discomfort connected to heavy resistance training (Fig 1). Baseline
characteristics (Tab 1) were measured by means of DXA (Prodigy Advance PA+302047, Lunar,
San Francisco, CA, USA) at Day -1, the last day preceding the RT intervention.

```{r DXA}
# Script: DXA

# Data handling
table.dxa <- read_excel("./data/dxa/ribose_dxa.xlsx") %>%
        select(sex, subject, age, height, weight, leanmasskg, fatmasskg, ffmg) %>%
        #make grams into kg
        mutate(ffkg = ffmg/1000) %>%
        select(sex, age, height, weight, leanmasskg, fatmasskg, ffkg) %>%
  mutate(bmi = (weight) / (height^2) * 10000)

dxa.table <- table.dxa %>%
        pivot_longer(names_to = "variable",
                     values_to = "values", col= age:bmi) %>%
        select(sex, variable, values) %>%
        group_by(sex, variable) %>%
        summarise(m = mean(values, na.rm = TRUE),
                  s= sd(values, na.rm = TRUE)) %>%
        mutate(stat = paste0(round(m, 1), " (", round(s, 1), ")")) %>%
        select(sex, variable, stat) %>%
        pivot_wider(names_from = sex,
                    values_from = stat) %>%
                mutate(variable = if_else(variable %in% c("age"),
                                          "Age (yrs)",
                                          if_else(variable %in% c("fatmasskg"),
                                                  "Fat mass (kg)",
                                                  if_else(variable %in% c("ffkg"),
                                                          "Fat free mass (kg)",
                                                          if_else(variable %in% c("height"),
                                                                  "Height (cm)",
                                                                  if_else(variable %in% c("leanmasskg"),
                                                                          "Lean body mass (kg)",
                                                                          if_else(variable %in% c("weight"),
                                                                                  "Body weight (kg)",
                                                                                  if_else(variable %in% c("bmi"),
                                                                                          "BMI", "")
                                                                                  )))))))

dxa.table %>%
        kable(col.names = c("Variable", "Female (n=7)", "Male (n=9)"))
```

#### Paper

##### Paper

###### Paper