### 2023-02-24 disp.Rmd
updates

---
title: "Disp"
output: word_document
date: "2023-02-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Intro

## Methods

### Materials and methods

### Participants

### Experimental design

### Dietary intervention

### Assessment of muscular strength

### Resistance training protocol

### Sampling of muscle tissue and blood

### Total RNA extraction and quantitative real-time reverse transcription polymerase chain reaction

### Protein extraction and immunoblotting

## Results

Baseline etc.
- Blood data
- Dietary data

### Markers of ribosome biogenesis?
- Total RNA? 
- Protein?
- UBF/total RNA?

### Muscular performance and recovery?

## Discussion