### 2023-02-28 results_work.RMD
fig text work

---
title: "results"
output: word_document
date: "2023-02-20"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r packages}
library(tidyverse);library(readxl);library(lme4);library(emmeans);library(dplyr);library(cowplot);library(knitr);library(broom);library(rmarkdown);library(tinytex);library(tidyr);library(arsenal);library(magrittr);library(dabestr);library(nlme); library(pbkrtest);library(ggpubr)
```

## Results

```{r glucose data + model}
## Scripts:
# - gluc.responsechange.R

# Data
glu.change <- readRDS("./data/data-gen/glucose/glu.change.RDS") # Untransformed data, used in figures
glu.logchange <- readRDS("./data/data-gen/glucose/glu.logchange.RDS") # Log-transformed data, used in the analysis

## Model
# Untransformed
glu.mod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
                     data = glu.change)
glu.res <- plot(glu.mod)
glu.sum <- summary(glu.mod)

# Log-transformed
glu.lmod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
                     data = glu.logchange)
glu.lres <- plot(glu.lmod)
glu.lsum <- summary(glu.lmod)

## Emmeans
# Untransformed
glu.emm <- confint(emmeans(glu.mod, specs = ~"supplement|time")) %>%
        data.frame()

# Log-transformed
glu.lemm <- confint(emmeans(glu.lmod, specs = ~"supplement|time")) %>%
        data.frame()

# For inline plotting
bcoef <- coef(summary(glu.lmod))
pscient.120 <- round(bcoef[8,5], 30)
p.120 <- sprintf("%.5f", 1.700971e-16)
pscient.135 <- round(bcoef[9,5], 30)
p.135 <- sprintf("%.5f", 4.299568e-11)
pscient.150 <- round(bcoef[10,5], 30)
p.150 <- sprintf("%.5f", 1.445712e-11)
p.270 <- round(bcoef[11,5], 2)
```

```{r c-peptide data + model}
## Scripts
# - insulin.change.R

## Data
cpep.change <- readRDS("./data/data-gen/glucose/cpep.change.RDS") # Untransformed data, for figures
cpep.lchange <- readRDS("./data/data-gen/glucose/cpep.logchange.RDS") # Log-transformed data, for analysis

## Linear mixed effects model
# Untransformed
cpep.mod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
                     data = cpep.change) 
cpep.res <- plot(cpep.mod)
cpep.sum <- summary(cpep.mod)

# Log-transformed
cpep.lmod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
                     data = cpep.lchange) 
cpep.lres <- plot(cpep.lmod)
cpep.lsum <- summary(cpep.lmod)

## Emmeans
# Untransformed
cpep.emm <- confint(emmeans(cpep.mod, specs = ~"supplement|time")) %>%
  data.frame()

# Log-transformed
cpep.lemm <- confint(emmeans(cpep.lmod, specs = ~"supplement|time")) %>%
  data.frame()

# For inline plotting
cpcoef <- coef(summary(cpep.lmod))
pscient.120 <- round(cpcoef[6,5], 30)
cpp.120 <- sprintf("%.5f", 3.229267e-15)
cpscient.150 <- round(cpcoef[7,5], 30)
cpp.150 <- sprintf("%.5f", 1.414491e-15)
cpp.270 <- round(cpcoef[8,5], 2)

```

At baseline, there were no significant differences between training with glucose or placebo in muscle mass, isometric or isokinetic peak torque, total session volume, training intensity, total RNA content, or rRNA expression (Tab 3). Further, on pairwise consecutive days, there were no significant differences in mean macro nutrient (calories, carbohydrate, fat or protein) ingestion (Tab 4), or mean change in macro nutrient ingestion (Fig 3). Glucose ingestion before and after RT led to significant increases in plasma levels of glucose and c-peptide compared to RT with placebo, before, during and after the session (Fig 1B, 120min: 38 ± 4% [*p* = `r p.120`], 135min: 31 ± 4% [*p* = `r p.135`], 150min: 32 ± 4% [*p* = `r p.150`], Fig 1C, 120min: 95 ± 10% [*p* = `r cpp.120`], 150min: 87 ± 10% [*p* = `r cpp.150`]), measured during the final RT session. At 2hrs post-exercise, RT with glucose led to significantly lower blood glucose levels compared to RT with placebo (1B, 270min: -8 ± 4% [*p* = `r p.270`].

FIG1 HERE

**Figure 1:** A) An overview of the experimental design with 12 days of concomitant dietary intervention and resistance training (RT), preceded by 7 days involving familiarization. Between days -7 and -1, participants were familiarized to the RT exercises via 1RM leg press and knee extension testing, and strength tests (Humac Norm dynamometer, days -7 and -5). Biopsies were taken from vastus lateralis at baseline (Day 1 leg 1, Day 2 leg 2), and after five RT sessions (Day 11 leg 1, Day 12 leg 2). Through randomization, one leg was allocated to RT with GLU (glucose) and the other leg to RT with PLA (placebo). The randomization also entailed whether participants initiated the intervention with their dominant or non-dominant leg, and whether they initiated with RT+GLU or RT+PLA. From baseline, the participants exercised in a unilateral, alternating fashion, until post intervention testing (Day 11/12 leg 1, Day 12/13 leg 2). Three participants dropped out of the intervention, either due to reported muscular discomforts of heavy RT, or sickness in their family. B and C) Changes in plasma glucose (mmol/L) and c-peptide levels (pmol/L) during post testing (Day 11 leg 1, Day 12 leg 2). Glucose levels in blood was measured via finger draws 120 (-120), 90 (-90), 30 (-30) minutes and immediately before RT (0), during RT (15), immediately after (30) and 2hrs after RT (120). C-peptide levels were measured simultaneously to these finger draws, except for 90min before and during RT. Finger draws and venous blood were sampled at baseline (Day 1) and post (Day 11 leg 1, Day 12 leg 2). D) Changes in muscular performance/recovery measured as isometric and isokinetic peak torque (60 and 240 d/s) via Humac Norm Dynamometer, conducted at baseline (Day -1), during the intervention (Day 4 and 8 leg 1, Day 5 and 9 leg 2), after five RT sessions/before the 6th session (Day 11 leg 1, Day 12 leg 2), as well as 30min, 2hrs and 23hrs after the 6th RT session (Day 11/12 leg 1, Day 12/13 leg 2). Values are presented as changes in estimated marginal means ± 95% CI. * = *p* < 0.05 between groups. Glucose n = 13, placebo n = 13.

### Markers of ribosome biogenesis

```{r total RNA model}
## Scripts
# - totrna.cleanup.R
# - totrna.analysis.R

# Data
trna.lchange <- readRDS("./data/data-gen/rna/totrna.change.RDS")

# Create model: 
trna.lmod <- lmerTest::lmer(change ~ pre + supplement + (1|subject), 
           data = trna.lchange)

trna.lres <- plot(trna.lmod)
trna.lsum <- summary(trna.lmod)

### Get estimated means from the model
trna.emm <- confint(emmeans(trna.lmod, specs = ~"supplement")) %>%
    data.frame()

# For inline reporting
trcoef <- coef(summary(trna.lmod))
trdiff <- round(trcoef[3,1], 2)
trdiffp <- round(trcoef[3,5], 2)
trdiffse <- round(trcoef[3,2], 2)

```

```{r rRNA data model}
## Scripts
# - qpcr-import
# - qpcr-analysis
# - rrna.change
# - rrnachange.figs

# Log-transformed data
lchange.18 <- readRDS("./data/data-gen/rna/lchange.18.RDS")
lchange.28 <- readRDS("./data/data-gen/rna/lchange.28.RDS")
lchange.58 <- readRDS("./data/data-gen/rna/lchange.58.RDS")
lchange.5 <- readRDS("./data/data-gen/rna/lchange.5.RDS")
lchange.47 <- readRDS("./data/data-gen/rna/lchange.47.RDS")

# Create model: 
# Needs to have an intercept per participant (mixed model)
# Control for pre values.

# 18S
lmod.18 <- lmerTest::lmer(change ~ pre + supplement + (1|subject), 
                     data = lchange.18)
lres.18 <- plot(lmod.18)
lsum.18 <- summary(lmod.18)

# 28S
lmod.28 <- lmerTest::lmer(change ~ pre + supplement + (1|subject), 
                     data = lchange.28)
lres.28 <- plot(lmod.28)
lsum.28 <- summary(lmod.28)

# 5.8S
lmod.58 <- lmerTest::lmer(change ~ pre + supplement + (1|subject), 
                     data = lchange.58)
lres.58 <- plot(lmod.58)
lsum.58 <- summary(lmod.58)

# 5S
lmod.5 <- lmerTest::lmer(change ~ pre + supplement + (1|subject), 
                     data = lchange.5)
lres.5 <- plot(lmod.5)
lsum.5 <- summary(lmod.5)

# 47S
lmod.47 <- lmerTest::lmer(change ~ pre + supplement + (1|subject), 
                     data = lchange.47)
lres.47 <- plot(lmod.47)
lsum.47 <- summary(lmod.47)

### Get estimated means from the model, these are average increase at 
# pre = 0 (the average pre value)
# remember that these are originally log-fold change values, transformed back using exp(emmean) etc

# 18S
lemm.18 <- confint(emmeans(lmod.18, specs = ~"supplement")) %>%
        data.frame()

# 28S
lemm.28 <- confint(emmeans(lmod.28, specs = ~"supplement")) %>%
        data.frame()

# 5.8S
lemm.58 <- confint(emmeans(lmod.58, specs = ~"supplement")) %>%
        data.frame()

# 5S
lemm.5 <- confint(emmeans(lmod.5, specs = ~"supplement")) %>%
        data.frame()

# 47S
lemm.47 <- confint(emmeans(lmod.47, specs = ~"supplement")) %>%
        data.frame()

## For inline reporting

# 18S
rr18coef <- coef(summary(lmod.18))
rr18diffp <- round(rr18coef[3,5], 3)
rr18diffse <- round(rr18coef[3,2], 3)

# 28S
rr28coef <- coef(summary(lmod.28))
rr28diffp <- round(rr28coef[3,5], 3)
rr28diffse <- round(rr28coef[3,2], 3)

# 5.8S
rr58coef <- coef(summary(lmod.58))
rr58diffp <- round(rr58coef[3,5], 3)
rr58diffse <- round(rr58coef[3,2], 3)

# 5S
rr5coef <- coef(summary(lmod.5))
rr5diffp <- round(rr5coef[3,5], 3)
rr5diffse <- round(rr5coef[3,2], 3)

# 47S
rr47coef <- coef(summary(lmod.47))
rr47diffp <- round(rr47coef[3,5], 3)
rr47tdiffp <- round(rr47coef[1,5], 3)
rr47diffse <- round(rr47coef[3,2], 3)

```

Resistance training with glucose did not induce a higher accumulation of total RNA (fig 2A, *p* = `r trdiffp`) or rRNA  (fig 2B, 47S: *p* = `r rr47diffp`, 18S: *p* = `r rr18diffp`, 28S: *p* = `r rr28diffp`, 5.8S: *p* = `r rr58diffp`, 5S: *p* = `r rr5diffp`) compared to RT with placebo, measured as mean change from baseline to post. From baseline to post-intervention, there was a mean increase in levels of total RNA by 26 and 22% after RT with glucose and placebo, respectively (*p* < 0.05, compared to baseline). A robust accumulation was also observed in ribosomal RNA expression in both RT with glucose and placebo, with mean increases between 34-43% (GLU) and 33-41% (PLAC) (*p* < 0.05, compared to baseline) in the four rRNA’s. The expression of in the 47S pre-rRNA increased by 37 and 59% with glucose and placebo respectively, where only RT with placebo increased significantly from baseline to post (fig 45, *p* = `r rr47tdiffp`).

FIG2 HERE

###### **Figure 2:** Changes in total RNA and ribosomal RNA. A) Total RNA, B) 47S pre-rRNA, 18S rRNA, 28S rRNA, 5.8S rRNA, 5S rRNA. Baseline = Day 1 leg 1/ Day 2 leg 2, Post = Day 11 leg 1, Day 12 leg 2. Total RNA and rRNA were analyzed in duplicates, with two duplicates per biopsy (two biopsies per time point), and normalized to ng x mg wet muscle weight for total RNA and external reference gene (Lambda) for rRNA. Total RNA and rRNA changes were calculated as log-fold change score per mg wet muscle weight. Mean change scores of the duplicates were calculated before modelling and transformed to the log-scale. Values are estimated marginal means fold change per leg per supplement ± 95% CI. Glucose n = 13, placebo n = 13.

```{r western modelling}
## Script: 
# - western.import.R
# - western.change.R

# Data

prot.dat <- readRDS("./data/data-gen/protein/prot.dat.RDS") # cleaned up and quantified western data
prot.out <- readRDS("./data/data-gen/protein/protein_complete.RDS") # Cleaned up and quantified western data, checked for outliers
prot.change <- readRDS("./data/data-gen/protein/prot.change.RDS") # change-score data

# By target

cmyc <- prot.change %>%
  filter(target == "cmyc")

ubf <- prot.change %>%
  filter(target == "ubf",
         subject != "103") # Subject 103 has missing values

rps6 <- prot.change %>%
  filter(target == "rps6")

# c-Myc change

m1 <- lme(change ~ pre + supplement,
            random = list(subject = ~ 1),
            data = cmyc) # same model as in "western import", but with change scores of norm.sign

#plot(m1)

sum.m1 <- summary(m1)

cmyc.emm <- emmeans(m1, specs = ~"supplement") %>%
  data.frame

# UBF change

m2 <- lme(change ~ pre + supplement,
            random = list(subject = ~ 1),
            data = ubf) # same model as in "western import", but with change (i.e., change in norm.sign) instead of norm.sign

#plot(m2)

sum.m2 <- summary(m2)

ubf.emm <- emmeans(m2, specs = ~"supplement") %>%
  data.frame()

# rps6 change

m3 <- lme(change ~ pre + supplement,
            random = list(subject = ~ 1),
            data = rps6) # same model as in "western import", but with change (i.e., change in norm.sign) instead of norm.sign

#plot(m3)

sum.m3 <- summary(m3)

rps6.emm <- emmeans(m3, specs = ~"supplement") %>%
  data.frame()

### For inline reporting
## Diff cmyc
ccoef <- coef(summary(m1))
cdiff <- round(ccoef[3,5], 3)
cdiffse <- round(ccoef[3,2], 3)

## Diff UBF
ucoef <- coef(summary(m2))
udiff <- round(ucoef[3,5], 3)
udiffse <- round(ucoef[3,2], 3)

## Diff rps6
rpcoef <- coef(summary(m3))
rpdiff <- round(rpcoef[3,5], 3)
rpdiffse <- round(rpcoef[3,2], 3)

```

```{r correlation model}
# Data 
west.dat <- readRDS("./data/data-gen/protein/prot.datc.RDS") # Quantified and cleaned western data,
                                                               # (outliers marked) normalised by pool

rna.dat <- readRDS("./data/data-gen/rna/rna.dat.RDS") # Quantified and cleaned (outliers marked), 
                                                      # total RNA data, normalised by wet muscle weight

## Prepare data frames for joining

ubf.rdy <- west.dat %>%  # One data frame without outliers
  select(subject, time, outlier, supplement, target, norm.sign) %>%
  filter(target == "ubf",
         outlier == "in") %>%
  select(-target) %>%
  group_by(subject, time, supplement) %>%
  summarise(mean.sign = mean(norm.sign, na.rm = TRUE)) %>%
  print()

rna.rdy <- rna.dat %>%
  select(subject, time, outlier, supplement, RNA.weight) %>%
  filter(outlier == "in") %>%
  mutate(subject = as.character(subject)) %>%
  group_by(subject, time, supplement) %>%
  summarise(mean.rna = mean(RNA.weight, na.rm = TRUE)) %>%
  print()

## Correlation analysis

joined.dat <- ubf.rdy %>%
  right_join(rna.rdy) %>%
  #filter(supplement != "GLUCOSE") %>%

  mutate(sd.ubf = (mean.sign - mean(mean.sign, na.rm = TRUE )/sd(mean.sign, na.rm = TRUE))) %>%
  print()

m <- lme(log(mean.rna) ~ mean.sign + time , 
         random = list(subject = ~ 1), 
          data = joined.dat, 
         na.action = na.omit)

summary(m)
plot(m)
```

Interestingly, RT with placebo led to significantly higher change in levels of rpS6 and UBF protein compared to RT with glucose, with no difference in levels of c-Myc (Fig 3A rpS6: *p* = `r rpdiff`, 3C UBF: *p* = `r udiff`, 3B c-Myc: *p* = `r cdiff`). Levels of rpS6 change by 40% in placebo and 7% in glucose (Fig 3A), UBF levels changed by 68% in placebo and 26% in glucose (Fig 3C), and c-Myc levels changed by 110% in placebo and 59% in glucose (Fig 3B).

FIG3 HERE

###### **Figure 3:** Changes in levels of protein. A) Representative total protein stain blot, B) AU fold change in UBF, c-Myc and rpS6 + representative antibody stain images of the respective proteins. Baseline = Training day 1/2, Post = Training day 11/12. Protein samples were analyzed in duplicates, with two duplicates per biopsy (two biopsies per time point). The duplicates were loaded on different gels, in an inverted order. Change in protein levels were calculated as log-fold change score normalised by pools (pools of all protein samples). Mean change scores of the duplicates were calculated before modelling and transformed to the log-scale. Values are estimated marginal means fold change per leg per supplement ± 95% CI. * = significant between groups statistic, glucose compared to placebo. Glucose n = 13, placebo n = 13.

Further, we observed a linear relationship between levels of UBF and accumulation of total RNA, where ...

FIG4 HERE

######**Figure 4:** Relationship between levels of total RNA and UBF. Total RNA was normalized by wet muscle weight, and UBF levels were normalized by pools and total protein per lane. Values are presented as log-transformed means.