### 2023-02-28 ext_method.RMD
added RMD for extended methods

---
title: "Ext. suppl."
author: "Kristian Lian"
date: "2023-02-28"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

Notes for extended methods supplementary

- UBF/RNA correlation: Make a new model to see if data point affects the current model