### 2023-03-01 results_work.RMD
updates before presentation

---
title: "results"
output: word_document
date: "2023-02-20"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r packages}
library(tidyverse);library(readxl);library(lme4);library(emmeans);library(dplyr);library(cowplot);library(knitr);library(broom);library(rmarkdown);library(tinytex);library(tidyr);library(arsenal);library(magrittr);library(dabestr);library(nlme); library(pbkrtest);library(ggpubr)
```

## Results

```{r glucose data + model}
## Scripts:
# - gluc.responsechange.R

# Data
glu.change <- readRDS("./data/data-gen/glucose/glu.change.RDS") # Untransformed data, used in figures
glu.logchange <- readRDS("./data/data-gen/glucose/glu.logchange.RDS") # Log-transformed data, used in the analysis

## Model
# Untransformed
glu.mod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
                     data = glu.change)
glu.res <- plot(glu.mod)
glu.sum <- summary(glu.mod)

# Log-transformed
glu.lmod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
                     data = glu.logchange)
glu.lres <- plot(glu.lmod)
glu.lsum <- summary(glu.lmod)

## Emmeans
# Untransformed
glu.emm <- confint(emmeans(glu.mod, specs = ~"supplement|time")) %>%
        data.frame()

# Log-transformed
glu.lemm <- confint(emmeans(glu.lmod, specs = ~"supplement|time")) %>%
        data.frame()

# For inline plotting
bcoef <- coef(summary(glu.lmod))
pscient.120 <- round(bcoef[8,5], 30)
p.120 <- sprintf("%.5f", 1.700971e-16)
pscient.135 <- round(bcoef[9,5], 30)
p.135 <- sprintf("%.5f", 4.299568e-11)
pscient.150 <- round(bcoef[10,5], 30)
p.150 <- sprintf("%.5f", 1.445712e-11)
p.270 <- round(bcoef[11,5], 2)
```

```{r c-peptide data + model}
## Scripts
# - insulin.change.R

## Data
cpep.change <- readRDS("./data/data-gen/glucose/cpep.change.RDS") # Untransformed data, for figures
cpep.lchange <- readRDS("./data/data-gen/glucose/cpep.logchange.RDS") # Log-transformed data, for analysis

## Linear mixed effects model
# Untransformed
cpep.mod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
                     data = cpep.change) 
cpep.res <- plot(cpep.mod)
cpep.sum <- summary(cpep.mod)

# Log-transformed
cpep.lmod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
                     data = cpep.lchange) 
cpep.lres <- plot(cpep.lmod)
cpep.lsum <- summary(cpep.lmod)

## Emmeans
# Untransformed
cpep.emm <- confint(emmeans(cpep.mod, specs = ~"supplement|time")) %>%
  data.frame()

# Log-transformed
cpep.lemm <- confint(emmeans(cpep.lmod, specs = ~"supplement|time")) %>%
  data.frame()

# For inline plotting
cpcoef <- coef(summary(cpep.lmod))
pscient.120 <- round(cpcoef[6,5], 30)
cpp.120 <- sprintf("%.5f", 3.229267e-15)
cpscient.150 <- round(cpcoef[7,5], 30)
cpp.150 <- sprintf("%.5f", 1.414491e-15)
cpp.270 <- round(cpcoef[8,5], 2)

```

```{r strength model}
## Scripts:
# - Strength.index.R

## Data
str.change <- readRDS("./data/data-gen/humac/str.change.RDS") # Untransfromed data
str.lchange <- readRDS("./data/data-gen/humac/str.lchange.RDS") # Log-transfromed data

## Linear mixed effects model

# Untransformed model
str.mod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
                         data = str.change)
str.res <- plot(str.mod)
str.sum <- summary(str.mod)

# Log-transformed model
str.lmod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
                        data = str.lchange)
str.lres <- plot(str.lmod)
str.lsum <- summary(str.lmod)

## Emmeans
# Untransformed
str.emm <- confint(emmeans(str.mod, specs = ~"supplement|time")) %>%
  data.frame()

# Log-transformed
str.lemm <- confint(emmeans(str.lmod, specs = ~"supplement|time")) %>%
  data.frame()

# For inline plotting
strcoef <- coef(summary(str.mod))
pdiff.c4 <- round(strcoef[10,5], 2)

```

At baseline, there were no significant differences between training with glucose or placebo in muscle mass, isometric or isokinetic peak torque, total session volume, training intensity, total RNA content, or rRNA expression (Tab 3). Further, on pairwise consecutive days, there were no significant differences in mean macro nutrient (calories, carbohydrate, fat or protein) intake, or mean change in macro nutrient ingestion (Tab 4).

Glucose ingestion before and after RT led to significant increases in plasma levels of glucose and c-peptide compared to RT with placebo, before, during and after the session (Fig 1B, 120min: 38 ± 4% [*p* = `r p.120`], 135min: 31 ± 4% [*p* = `r p.135`], 150min: 32 ± 4% [*p* = `r p.150`], Fig 1C, 120min: 95 ± 10% [*p* = `r cpp.120`], 150min: 87 ± 10% [*p* = `r cpp.150`]), measured during the final RT session. At 2hrs post-exercise, RT with glucose led to significantly lower blood glucose levels compared to RT with placebo (1B, 270min: -8 ± 4% [*p* = `r p.270`]. There were no significant difference in change of skeletal muscle strength, at time points post 2 RT sessions, post 4 RT sessions, 30min after 6th session, 2hrs after 6th session and 23hrs after 6th session (*p* > 0.05). After 5 RT sessions (Fig 1D, Post 5RT), there was a significant difference in strength between glucose and placebo,  where placebo decreased strength 7% more than glucose (*p* = `r pdiff.c4`).

**Figure 1:** A) An overview of the experimental design with 12 days of concomitant dietary intervention and resistance training (RT), preceded by 7 days involving familiarization. Between days -7 and -1, participants were familiarized to the RT exercises via 1RM leg press and knee extension testing, and to the strength tests via Humac Norm dynamometer (days -7 and -5). Before baseline testing, the participants were randomly allocated to exercise one leg with GLU (glucose) and the other with PLA (placebo), in a unilateral, alternating fashion. Further, non-dominant/dominant + GLU/PLA, and onset with GLU or PLA was also randomized, i.e. the figure illustrates an example where the participant was randomized to start RT with GLU. Biopsies were taken from m. vastus lateralis at baseline (Day 1 leg 1, Day 2 leg 2), and after five RT sessions (Day 11 leg 1, Day 12 leg 2). Three participants dropped out of the intervention, either due to reported muscular discomforts of heavy RT, or sickness in their family. B and C) Changes in plasma glucose (mmol/L) and c-peptide levels (pmol/L) during post testing (Day 11 leg 1, Day 12 leg 2). Glucose levels in blood was measured via finger draws 120 (-120), 90 (-90), 30 (-30) minutes and immediately before RT (0), during RT (15), immediately after (30) and 2hrs after RT (120). C-peptide levels were measured simultaneously to these finger draws, except for 90min before and during RT. Finger draws and venous blood were sampled at baseline (Day 1) and post (Day 11 leg 1, Day 12 leg 2). D) Changes in muscular performance/recovery measured as isometric and isokinetic peak torque (60 and 240 d/s) via Humac Norm Dynamometer, conducted at baseline (A: Day -1), after 2 and 4 RT sessions (A: Day 4 and 8 leg 1, Day 5 and 9 leg 2), after five RT sessions/before the 6th session (Day 11 leg 1, Day 12 leg 2), as well as 30min, 2hrs and 23hrs after the 6th RT session (A: Day 11/12 leg 1, Day 12/13 leg 2). The index was calculated by normalizing individual peak torque to the highest peak torque at each respective speed, and then summarised and used in change score calculations.

Values are presented as changes in estimated marginal means ± 95% CI. * = *p* < 0.05 between groups. Glucose n = 13, placebo n = 13.

### Markers of ribosome biogenesis

```{r total RNA model}
## Scripts
# - totrna.cleanup.R
# - totrna.analysis.R

# Data
trna.lchange <- readRDS("./data/data-gen/rna/totrna.change.RDS")

# Create model: 
trna.lmod <- lmerTest::lmer(change ~ pre + supplement + (1|subject), 
           data = trna.lchange)

trna.lres <- plot(trna.lmod)
trna.lsum <- summary(trna.lmod)

### Get estimated means from the model
trna.emm <- confint(emmeans(trna.lmod, specs = ~"supplement")) %>%
    data.frame()

# For inline reporting
trcoef <- coef(summary(trna.lmod))
trdiff <- round(trcoef[3,1], 2)
trdiffp <- round(trcoef[3,5], 2)
trdiffse <- round(trcoef[3,2], 2)

```

```{r rRNA data model}
## Scripts
# - qpcr-import
# - qpcr-analysis
# - rrna.change
# - rrnachange.figs

# Log-transformed data
lchange.18 <- readRDS("./data/data-gen/rna/lchange.18.RDS")
lchange.28 <- readRDS("./data/data-gen/rna/lchange.28.RDS")
lchange.58 <- readRDS("./data/data-gen/rna/lchange.58.RDS")
lchange.5 <- readRDS("./data/data-gen/rna/lchange.5.RDS")
lchange.47 <- readRDS("./data/data-gen/rna/lchange.47.RDS")

# Create model: 
# Needs to have an intercept per participant (mixed model)
# Control for pre values.

# 18S
lmod.18 <- lmerTest::lmer(change ~ pre + supplement + (1|subject), 
                     data = lchange.18)
lres.18 <- plot(lmod.18)
lsum.18 <- summary(lmod.18)

# 28S
lmod.28 <- lmerTest::lmer(change ~ pre + supplement + (1|subject), 
                     data = lchange.28)
lres.28 <- plot(lmod.28)
lsum.28 <- summary(lmod.28)

# 5.8S
lmod.58 <- lmerTest::lmer(change ~ pre + supplement + (1|subject), 
                     data = lchange.58)
lres.58 <- plot(lmod.58)
lsum.58 <- summary(lmod.58)

# 5S
lmod.5 <- lmerTest::lmer(change ~ pre + supplement + (1|subject), 
                     data = lchange.5)
lres.5 <- plot(lmod.5)
lsum.5 <- summary(lmod.5)

# 47S
lmod.47 <- lmerTest::lmer(change ~ pre + supplement + (1|subject), 
                     data = lchange.47)
lres.47 <- plot(lmod.47)
lsum.47 <- summary(lmod.47)

### Get estimated means from the model, these are average increase at 
# pre = 0 (the average pre value)
# remember that these are originally log-fold change values, transformed back using exp(emmean) etc

# 18S
lemm.18 <- confint(emmeans(lmod.18, specs = ~"supplement")) %>%
        data.frame()

# 28S
lemm.28 <- confint(emmeans(lmod.28, specs = ~"supplement")) %>%
        data.frame()

# 5.8S
lemm.58 <- confint(emmeans(lmod.58, specs = ~"supplement")) %>%
        data.frame()

# 5S
lemm.5 <- confint(emmeans(lmod.5, specs = ~"supplement")) %>%
        data.frame()

# 47S
lemm.47 <- confint(emmeans(lmod.47, specs = ~"supplement")) %>%
        data.frame()

## For inline reporting

# 18S
rr18coef <- coef(summary(lmod.18))
rr18diffp <- round(rr18coef[3,5], 3)
rr18diffse <- round(rr18coef[3,2], 3)

# 28S
rr28coef <- coef(summary(lmod.28))
rr28diffp <- round(rr28coef[3,5], 3)
rr28diffse <- round(rr28coef[3,2], 3)

# 5.8S
rr58coef <- coef(summary(lmod.58))
rr58diffp <- round(rr58coef[3,5], 3)
rr58diffse <- round(rr58coef[3,2], 3)

# 5S
rr5coef <- coef(summary(lmod.5))
rr5diffp <- round(rr5coef[3,5], 3)
rr5diffse <- round(rr5coef[3,2], 3)

# 47S
rr47coef <- coef(summary(lmod.47))
rr47diffp <- round(rr47coef[3,5], 3)
rr47tdiffp <- round(rr47coef[1,5], 3)
rr47diffse <- round(rr47coef[3,2], 3)

```

Resistance training with glucose did not induce a higher accumulation of total RNA (fig 2A, *p* = `r trdiffp`) or rRNA  (fig 2B, 47S: *p* = `r rr47diffp`, 18S: *p* = `r rr18diffp`, 28S: *p* = `r rr28diffp`, 5.8S: *p* = `r rr58diffp`, 5S: *p* = `r rr5diffp`) compared to RT with placebo, measured as mean change from baseline to post. From baseline to post-intervention, there was a mean increase in levels of total RNA by 26 and 22% after RT with glucose and placebo, respectively (*p* < 0.05, compared to baseline). A robust accumulation was also observed in ribosomal RNA expression in both RT with glucose and placebo, with mean increases between 34-43% (GLU) and 33-41% (PLAC) (*p* < 0.05, compared to baseline) in the four rRNA’s. The expression of in the 47S pre-rRNA increased by 37 and 59% with glucose and placebo respectively, where only RT with placebo increased significantly from baseline to post (fig 45, *p* = `r rr47tdiffp`).

FIG2 HERE

###### **Figure 2:** Changes in total RNA and ribosomal RNA. A) Total RNA, B) 47S pre-rRNA, 18S rRNA, 28S rRNA, 5.8S rRNA, 5S rRNA. Baseline = Day 1 leg 1/ Day 2 leg 2, Post = Day 11 leg 1, Day 12 leg 2. Total RNA and rRNA were analyzed in duplicates, with two duplicates per biopsy (two biopsies per time point), and normalized to ng x mg wet muscle weight for total RNA and external reference gene (Lambda) for rRNA. Total RNA and rRNA changes were calculated as log-fold change score per mg wet muscle weight. Mean change scores of the duplicates were calculated before modelling and transformed to the log-scale. Values are estimated marginal means fold change per leg per supplement ± 95% CI. Glucose n = 13, placebo n = 13.

```{r western modelling}
## Script: 
# - western.import.R
# - western.cleanup.R
# - western.analysis.R

# Data
prot.dat <- readRDS("./data/data-gen/protein/prot.datc.RDS") # from the script: western.cleanup.R

### Modelling target signal per supplement 
# These models try to explain the normalised signal by time, supplement and time+supplement (?). The random = list command creates an individual 
# intercept for the included (e.g., subject, gel.sample).

## c-Myc model

# Untransformed
cmyc.mod <- lme(norm.sign ~ time + supplement + time:supplement, 
                random = list(subject = ~ 1),
                data = filter(prot.dat, target == "cmyc"))

cmyc.sum <- summary(cmyc.mod)
cmyc.res <- plot(cmyc.mod, resid(., type = "p") ~ fitted(.))

# Log-transformed
cmyc.lmod <- lme(log(norm.sign) ~ time + supplement + time:supplement, 
                 random = list(subject = ~ 1),
                 data = filter(prot.dat, target == "cmyc"))

cmyc.lsum <- summary(cmyc.lmod)
cmyc.lres <- plot(cmyc.lmod, resid(., type = "p") ~ fitted(.))

## UBF model
# Untransformed
ubf.mod <- lme(norm.sign ~ time + supplement + time:supplement, 
               random = list(subject = ~ 1),
               data = filter(prot.dat, target == "ubf"))

ubf.sum <- summary(ubf.mod)
ubf.res <- plot(ubf.mod, resid(., type = "p") ~ fitted(.))

# Log-transformed
ubf.lmod <- lme(log(norm.sign) ~ time + supplement + time:supplement, 
                random = list(subject = ~ 1),
                data = filter(prot.dat, target == "ubf"))

ubf.lsum <- summary(ubf.lmod)
ubf.lres <- plot(ubf.lmod, resid(., type = "p") ~ fitted(.))

# RPS6 model
# Untransformed
rps6.mod <- lme(norm.sign ~ time + supplement + time:supplement, 
                 random = list(subject = ~ 1),
                 data = filter(prot.dat, target == "rps6"))

rps6.sum <- summary(rps6.mod)
rps6.res <- plot(rps6.mod, resid(., type = "p") ~ fitted(.))

# Log-transformed
rps6.lmod <- lme(log(norm.sign) ~ time + supplement + time:supplement, 
          random = list(subject = ~ 1),
          data = filter(prot.dat, target == "rps6"))

rps6.lsum <- summary(rps6.lmod)
rps6.lres <- plot(rps6.lmod, resid(., type = "p") ~ fitted(.))

### Getting emmeans from the models

## c-Myc emmeans
# Untransformed
cmyc.emm <- emmeans(cmyc.mod, specs = ~ time|supplement) %>%
  data.frame()

# Log-transformed
cmyc.lemm <- emmeans(cmyc.lmod, specs = ~ time|supplement) %>%
  data.frame()

## UBF emmeans
# Untransformed
ubf.emm <- emmeans(ubf.mod, specs = ~ time|supplement) %>%
  data.frame()

# Log-transformed
ubf.lemm <- emmeans(ubf.lmod, specs = ~ time|supplement) %>%
  data.frame()

## RPS6
# Untransformed
rps6.emm <- emmeans(rps6.mod, specs = ~ time|supplement) %>%
  data.frame()

# Log-transformed
rps6.lemm <- emmeans(rps6.lmod, specs = ~ time|supplement) %>%
  data.frame()

### For inline reporting
## Diff cmyc

ccoef <- coef(summary(cmyc.lmod))
cdiff <- round(ccoef[4,5], 3)
cpla <- round(ccoef[2,5], 20)

## Diff UBF
ucoef <- coef(summary(ubf.lmod))
udiff <- round(ucoef[4,5], 3)

## Diff rps6
rpcoef <- coef(summary(rps6.lmod))
rpdiff <- round(rpcoef[4,5], 3)

```

```{r correlation model}
# Data 
west.dat <- readRDS("./data/data-gen/protein/prot.datc.RDS") # Quantified and cleaned western data,
                                                               # (outliers marked) normalised by pool

rna.dat <- readRDS("./data/data-gen/rna/rna.dat.RDS") # Quantified and cleaned (outliers marked), 
                                                      # total RNA data, normalised by wet muscle weight

## Prepare data frames for joining

ubf.rdy <- west.dat %>%  # One data frame without outliers
  select(subject, time, outlier, supplement, target, norm.sign) %>%
  filter(target == "ubf",
         outlier == "in") %>%
  select(-target) %>%
  group_by(subject, time, supplement) %>%
  summarise(mean.sign = mean(norm.sign, na.rm = TRUE)) %>%
  print()

rna.rdy <- rna.dat %>%
  select(subject, time, outlier, supplement, RNA.weight) %>%
  filter(outlier == "in") %>%
  mutate(subject = as.character(subject)) %>%
  group_by(subject, time, supplement) %>%
  summarise(mean.rna = mean(RNA.weight, na.rm = TRUE)) %>%
  print()

## Correlation analysis

joined.dat <- ubf.rdy %>%
  right_join(rna.rdy) %>%
  #filter(supplement != "GLUCOSE") %>%

  mutate(sd.ubf = (mean.sign - mean(mean.sign, na.rm = TRUE )/sd(mean.sign, na.rm = TRUE))) %>%
  print()

library(nlme)

corr.mod <- lme(mean.rna ~ mean.sign + time , 
         random = list(subject = ~ 1), 
          data = joined.dat, 
         na.action = na.omit)

corr.sum <- summary(corr.mod)
corr.res <- plot(corr.mod)

corr.lmod <- lme(log(mean.rna) ~ log(mean.sign) + time , 
         random = list(subject = ~ 1), 
          data = joined.dat, 
         na.action = na.omit)

corr.lsum <- summary(corr.lmod)
corr.lres <- plot(corr.lmod)

corrcoef <- coef(summary(corr.mod))
corrp <- round(corrcoef[2,5], 3)
```

C-myc levels increased 109% after RT with placebo, compared to an increase of 58% after RT with glucose, which was a significant difference (Fig 3A: *p* = `r cdiff`). Levels of UBF and rpS6 increased 42 and 33% more in placebo compared to glucose, however, this was not a significant difference (Fig 3A: UBF: *p* = `r udiff`, rpS6: *p* = `r rpdiff`). Compared to baseline, both RT with glucose and RT with placebo increased significantly () Further, we observed a linear relationship between levels of UBF and accumulation of total RNA, where an increase of 1 SD unit of UBF equated to 13% increase in total RNA (*p* = `r corrp`).

FIG3 HERE

###### **Figure 3:** Changes in levels of protein.

A) Mean UBF, c-Myc and rpS6 (AU) at baseline and post + representative antibody stains of the respective proteins. Baseline = Day 1 leg 1, Day 2 leg 2, post = Day 11 leg 1, Day 12 leg 2. GLU = glucose, PLA = placebo. Protein samples were analyzed in two duplicates per biopsy per time point, loaded on separate gels in an inverted order, e.g. from gel 3 to 6, as shown by the duplicates (1 and 2). Change in protein levels were calculated as log-fold change scores normalized by pools (pool of all protein samples per gel). Mean change scores of the duplicates were calculated before modelling and transformed to the log-scale. Values are reverse transformed estimated marginal means fold change per leg per supplement ± 95% CI. * = significant between groups statistic, glucose compared to placebo. Glucose n = 13, placebo n = 13. B) Representative total protein stain blot. C) Linear relationship between total RNA (ng x mg) and UBF levels (SD units), with time added as a covariate. Total RNA was normalized by wet muscle weight, and UBF was normalized by pools per gel, and total protein per lane factor. Values are presented as log-transformed means.

#MISSING: CHECK % INCREASE OF TOTAL RNA PER 1 SD INCREASE OF UBF (MAY HAVE TO LOG-TRANSFORM)

(We must add time as a covariate, as both total RNA and UBF are expected to increase with "time" in this intervention, due to RT. Therefore, adding time as a covariate allows us to model the relationship between RNA and UBF, without variations of time interfering)