### 2023-03-07 disp.Rmd
updates

---
title: "Disp"
output: word_document
date: "2023-02-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Intro
- To do: Check remaining comments

## Methods

### Materials and methods

### Participants

### Experimental design

### Dietary intervention

### Assessment of muscular strength

### Resistance training protocol

### Sampling of muscle tissue and blood

### Total RNA extraction and quantitative real-time reverse transcription polymerase chain reaction

### Protein extraction and immunoblotting

- To do: Read through - anything needs changing?

## Results

Baseline etc.
- Fig1 + tabs?
- Dietary data comments

### Markers of ribosome biogenesis?

#### RNA and rRNA
- Fig2

#### Protein
- Fig3

- To do: Check figure text, implement tabs and figures

## Discussion

Main topics:
- No difference between glu and plac in main outcomes - thoughts?
  - No additional effect of glu, as long as daily macro intake is equal
  - No direct effect of glu on anabolic signalling, i.e., glucose per se does not seem to induce RT responses 
  - Based on findings in litterature, likely still important as a contributor to the cells energy pool, thus indirectly
    affecting ribosome biogenesis

- Is there any coherence between strength diff post session 5 and cmyc diff?
  - I.e., post 5 data indicate that placebo experiences more fatigue, does this cause higher c-Myc levels, since c-Myc is a           general transcription factor responding to stress?
  - There were no differences in total session volumes between glucose and placebo - how could placebo then experience more stress     than glucose?
  - Does glucose, e.g. via increasing energy availability, reduce metabolic stress? If so, is that a good thing for adaptations?      If there is less stress, is there less stimulation to adapt to? Mechanical strain vs metabolic stress? 
  - "Bifacial recovery"?

- UBF/RNA: 
  - Ref to Hammarström 21, similar observations
  - One of the first studies to find this relationship in vivo in human muscle cells 
  - The same was not true for c-Myc and rpS6, further strengthening the importance of UBF in ribosome biogenesis compared to c-Myc
(We must add time as a covariate, as both total RNA and UBF are expected to increase with "time" in this intervention, due to RT. Therefore, adding time as a covariate allows us to model the relationship between RNA and UBF, without variations of time interfering)

- Any perspectives to pick up from the thesis?