### 2023-03-10 ext_method.RMD
updated methods and results in ribose-paper.RMD

---
title: "Ext. suppl."
author: "Kristian Lian"
date: "2023-02-28"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

Notes for extended methods supplementary

- Primer sequence table?
- Antiboy table?
- Power calculation + rational?
- Normalization of total protein, qpcr values and protein signals
- UBF/RNA correlation: Make a new model to see if data point affects the current model

Mean-centered baseline values were used to correct for regression to the mean, which potentially occurs when the same participants
are repeatedly tested. Regression to the mean entails that scores close to the upper or lower limit of a participants potential score would be replaced with a score closer to the mean, e.g. high baseline values are shifted downwards while low baseline values are shifted upwards, creating a negative association between initial values and change. Estimated marginal means were calculated from the linear model, enabling acquisition of least-square means, showing the means for all involved groups adjusted to means of other potential factors in the model.

## Code/formulas for quantitative analyses
Total RNA: 
- Data import and prepping
 - Packages: tidyverse, readxl, nlme
 - Script: totrna.cleanup.R
 - Raw data = protein concentration and 260/280 ratio from excel sheet "RNA_raw"
 - Checked for outliers via mutate(resid = resid/sd(resid), outlier = if_else(resid < -2,      "out", "in"))
  - This code first calculates the samples residuals, then creates a variable called "outlier", containing "in" and "out".
    Further, it marks samples with "out" if the residuals are greater than 2, and "in" if residuals are lesser than 2. Thus, 
    we can choose to filter out participants marked with out

- Change score calculation
 - Packages: tidyverse
 - Script: totrna.analysis.R
 - Before change score calculation, total RNA is normalized by wet muscle weight
  - Approach: mean RNA / mean weight = rna.weight. Then, mean rna.weight is used to calculate change scores
 - Cleaned up data was used to calculated the change score from pre to post
  - Visual inspection of the residuals resulted in log-transforming the data, to control for heteroscedasticity
  - Calculation: change = log(post)-log(pre), pre = pre - mean(pre, na.rm = TRUE).
   - pre = pre - mean(pre, na.rm = TRUE): this code subtracts the mean rna concentration at pre from each subjects mean,
     so we can mean-center the baseline values to control for regression to the mean

- Analysis:
 - Packages: tidyverse, lme4, nlme, emmeans, readxl
 - Script: totrna.analysis.R
 - Model: lmerTest::lmer(change ~ pre + supplement + (1|subject), data = change_dat)
  - In this model, change is added as the response, with pre (baseline values) and supplement (treatment, glucose/placebo) as         explanatory factors
  - (1|subject) creates an intercept for each participant? why?
 - Emmeans: estimated marginal means, allows calculation of least square means - why?

rRNA
- Data import and prepping
 - Packages: qpcr, qpcrpal, parallell, tidyverse
 - Script: qpcr.import.R
 - Imports raw fluorescence data and models amplification curves with a best-fit sigmoidal model using the qPCR package
 - Estimates primer efficiency and cq-values

- Change score calculation
 - Packages: tidyverse
 - Script: qpcr.analysis.R
 - Before change score calculation, qpcr-values are normalized by wet muscle weight and external reference gene (Lambda)
  - ..
- Analysis
 - Analyses were done with the same approach and model as for total RNA, with normalize cq-values per target 
  - 47S, 18S, 28S, 5.8S and 5S

Protein
- Data import and prepping
 - Packages: tidyverse, readxl
 - Script: western.import.R, western.cleanup.R
 - Normalization:
  - We wanted to normalize signal data to total protein per gel and pools, to be able to        compare gels and protein samples from with varying amount of total protein
  - First, we quantified mean gray from within and between the lanes of each gel in ...         (software).
  - Mean gray within lanes were categorized as total protein, mean gray between lanes were      categorized as background and summarised per gel
  - Then we subtracted background from total protein, creating the total protein per lane       (tpl) factor
  - To make this factor comparable across gels, we divided it by the maximal tpl per gel        (tpl = tpl/max(tpl))
  - We also did this for the mean protein signal (signal = signal/max(signal))
  - Then, we divided signal by tpl, creating the norm.sign factor (norm.sign = signal/tpl).
  - norm.sign was lastly divided by mean signal from the pooled samples (norm.sign =            norm.sign/pool), resulting in protein signals comparable across different gels and with     varying amount of protein sample (ng/ul)
 - Clean up
  - Quantified western data was cleaned up the same way as total RNA, with marking samples 
    above and below 2 residuals
- Analysis
 - The model used for the protein data was from the lme4 package
 - model: lme(log(norm.sign) ~ time + supplement + time:supplement, 
          random = list(subject = ~ 1),
          data = filter(prot.dat, target == "rps6"))
  - Here, norm.sign (i.e. normalized mean protein signal) is the response, while time           (baseline and post) + supplement (treatment) and time: supplement are the explanatory       factor/terms
  - the model uses a two-sided linear formula to described the fixed effects (og(norm.sign)     ~ time + supplement + time:supplement)
   - This entails describing whether there are differences between pre and post (time),          between treatments (supplement), and both?
   - random effects (random = list(subject = ~ 1))?
   - Why did we use a different model?

UBF/Total RNA
- Data prepping
- Packages: tidyverse, ggpubr
- Scripts: ubf_rna.R
 - For this analysis, we used cleaned up data from total RNA and western blot analyses
 - UBF was filtered out from the clean western data
 - mean.ubf and mean.rna were summarised in their respective data frames, then joined for 
   the model
  - Further, we calculated the SD units of ubf to model with total rna, because ...
   - calculation: mutate(sd.ubf = (mean.sign - mean(mean.sign, na.rm = TRUE )/sd(mean.sign,      na.rm = TRUE)))
- Analysis
 - this model was built with the nlme package
 - model: lme(mean.rna ~ mean.sign + time , 
              random = list(subject = ~ 1), 
              data = joined.dat, 
              na.action = na.omit)
 - Here the mean.rna is the response, with the terms being mean.sign and time
   we added time as a covariate here, since both levels of ubf and total rna are expected
   to vary with time
 - random effect (random = list(subject = ~ 1)) - why ?

Strenght index
- Data import and cleanup
 - Strength index was calculated with peak torque values from all humac tests
  - I.e. baseline -> post session 60, all speeds
  - To manage this, we first calculated an index for each test speed (0, 60 and 240 d/s), by     by grouping peak torque values by time and supplement, and diving it by the maximum peak     torqu.
  - Then, we summarised the peak torque indexes by subject, time and supplement and gathered     all the values under the variable "strenght index".
  - This way, we can compare peak torque values from different movement speeds, as the index     puts them on a comparable scale
- Change score calculations
 - change score was done similarly to total RNA, across multiple time points
- Analysis
 - To account for the multiple time points, the model was changed accordingly
 - Model: lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
                         data = str.change)
 - Why do we add the zero?
 - In this model, we have both baseline and time as covariates, to account for the multiple    time points

Plasma glucose and c-peptide
- Data import and prepping
 - Data was imported and prepped in the same way as previously described data,
   by summarising means at each time point
- Change score calculation
 - change scores was calculated in the same way as previously described
- Analysis
 - The same model as strength index