### 2023-03-21 ext_method.RMD
updated ext_method

---
title: "Ext. suppl."
author: "Kristian Lian"
date: "2023-02-28"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

Notes for extended methods supplementary

- Antiboy table?
- Power calculation + rational?

## Wet lab analyses

We gathered muscle tissue from both legs at pre and post intervention, taking two samples from the same incision. Each piece of muscle tissue was split in two, i.e. we had two samples for RNA analyses and two samples for protein analyses. We analyzed the samples in duplicates, i.e. we had two samples per time point per leg, a total of eight samples per analysis. <!-- check if there is more tissue left --> These samples were then randomly labelled with participant number and a number between 1-8, to blind the investigators conducting the laboratory analyses.

### Total RNA and rRNA

#### Extraction of RNA from muscle tissue (protocol)

##### Materials

For extraction of RNA and protein per sample:

1.5 ml Eppendorf tube with 0.5 mm RNase-free Zirconium beads (~50 ul) Keep out DNA RNAse contaminations (Eppendorf safe-lock for use in bullet blender.
1.5 ml tube for RNA precipitation
1.5 ml tube for RNA quantification (1:5 dilution, add 20 ul of DEPC-water)
1.5 ml tubes for Protein precipitation
1.5 ml tubes for Soluble protein fraction
1.5 ml tubes for protein quantification (1:10 dilution, add 36 ul of ddH2O)

##### Solutions and chemicals

Trizol (Life-technologies)
Chloroform
Isopropanol
External reference gene (e.g. Takara lambda)
100% EtOH
95% EtOH
75% EtOH (kept on -20)

##### Homogenization and phase separation

1. Add 1 sample volume of beads and 300 ul Trizol to eppendorf 1.5 ml tubes. Keep on ice. OPTIONAL: If external reference gene is used, add 1 unit to every extraction with Trizol in first step. Lambda (0.04 ng/ul, 2 ul per extraction)
2. Add frozen muscle samples (protocol optimized for 5-25 ug of wet weight muscle sample) without thawing and put tubes in Bullet blender
3. Run Bullet Blender (BB) for 1 min at speed 10, put samples on ice for 1 min. Run samples for 1 min at speed 12. Check samples, if not dissolved run until fully homogenized.
4. Add Trizol (700 ul) to total volume 1000 ul, run samples in BB for 1 min speed 3.
5. Incubate @ RT for 5 min
6. Add 200 ul chloroform, shake 15 sec, incubate for 2-3 min.
7. Spin 12000g, 15min, 4°C
8. Pipette 450 ul of the aqueous phase to a new tube. Be careful not to disrupt the interphase or organic phase. Add 500 ul of isopropanol to precipitate RNA, incubate over night at -20°C or 10 min at room temp and continue with RNA extraction.
9. Completely remove the remaining aqueous phase, either store remaining phases at -20°C until protein extraction or continue right away

##### Extraction

Precipitation and wash

1. Spin the isopropanol/water phase 12000g, 10min, 4°C. A pellet should form, decant supernatant, spin briefly and remove remaining liquids.
2. Add 1000 ul 75% EtOH (-20°C) @ RT, vortex gently. Spin 7500g, 5min, 4°C.
3. Repeat step 2 two more times.
4. Remove all EtOH, air dry the pellet up to 10 min with tubes upside-down. Do not over-dry!

Suspension
1. Add 20-40 ul of 1x DEPC-treated water (55°C), incubate at 55°C for 10min.
2. Aliquot:
 5 ul to 20 ul TE buffer for Nanodrop/Qubit2
 x ul in stock solution
3. Store RNA stock solution @ -80°C.

#### cDNA synthesis using Superscript IV and general qPCR

##### cDNA

##### Materials

Heating block or thermal-cycler capable of 65, 50-55 and 80°C
RNase-free tubes, pipette tips
Superscript IV kit (SSIV Buffer, SSIV Reverese transcriptase, 100 nM DTT)
Oligo dT and/or Random Hexamer primers
10 mM dNTP mix
Total-RNA or mRNA

1. Combine the following components per reaction (consider creating a master mix from component a and b if you are running many samples, master mix 1):
 25 μM Oligo dT and 25 μM Random Hexamers - 0.5 μl Oligo dT (art nr. AB-1247) + 0.5 μl Random hexamers (SO142) per reaction
 10 mM dNTP mix (10 mM each) - 1 μl per reaction
 Template RNA (0.1-5000 ng total-RNA) - up to 11 μl per reaction
 Nuclease-free water - Add to total volume 13 μl per reaction

2. Mix all components, vortex and briefly spin down.
3. Heat the mix 5 min at 65°C followed by at least 1 min incubation on ice.
4. Bring the 5x SSIV buffer to room-temp. Vortex and spin briefly.
5. Combine the following into master-mix 2:
 5x SSIV Buffer - 4 μl per reaction
 100 mM DTT - 1 μl per reaction
 RNase OUT - 1 μl per reaction
 Super Script reverse transcriptase - 1 μl per reaction

6. Vortex and briefly spin the components of master mix 2, add master-mix 2 to the reaction tube.
7. Incubate the reaction tube at 23°C 10 min.
8. Incubate the reaction mix at 50-55 °C 10 min followed by 10 min at 80°C
9. Dilute the cDNA to desired concentration (e.g. 1:25)

##### qPCR

##### Materials

QuantStudio 5 real-time PCR machine
A qPCR reaction plate
Nuclease-free water and pipette tips
SYBR-green Master mix

1. Combine master-mix:
 Sybre-green 2X master-mix - 5 μl per reaction
 Primermix, Forward and Reverse 5 μM each - 1 μl per reaction
 H2O - 2 μl per reaction

2. Load the plate with primer-specific master-mix
3. Add 2 μl cDNA sample

### Protein

#### Extraction

###### **Table 3:** Primer sequence and performance. Average cycle thresholds (Ct) and primer efficiencies were calculated from all qPCR reactions.

![](figures/archive/Primer.sequences.PNG)

## Statistical analyses

We chose to log-transformed all data, as visual inspection showed clear signs of heteroscedasticity in some of the data sets. Thus, data was log-transformed prior to modelling with the function log(), and then reverse-transformed for figure illustration via the function exp(). Exceptions to this was the data used in figure 1 (B, C and D). In figure 1B and -C, the illustrated data are absolute changes in plasma glucose and c-peptide, as this demonstrated quite well the change that occured, and arguably, is quite more understandable compared to arbitrary units, such as the protein signal data. For figure 1D, we had already made in index of strength combining isometric and isokinetic peak torque, meaning the data was already put on a different scale than the original absolute change in peak torque.

Further, we mean-centered the baseline values in each model to account for regression to the mean. This was simply done by subtracting the mean baseline value from each participants baseline value, to control for the possibility of extremely high or low baseline values to violate the assumption of equal variance. Estimated marginal means were calculated from the linear models, enabling acquisition of least-square means, showing the means for all involved groups adjusted to means of other variables in the model. This is a way to statistically model the means of the data, rather than describing them as they appear in the data.

To enable cross nano Drop, qPCR plate and gel comparisons from the wet lab analyses, we normalized RNA and protein data. Total RNA was normalized to wet muscle weight, rRNA was normalized to wet muscle weight and the external reference gene Lambda, while protein data was normalized both to the total protein lane factor and to pools of all protein samples.

RNA normalization: 
 1. weight = mean(weight) 
 2. RNA = mean(RNA)
 3. weight.mc = weight/mean(weight)
 4. RNA.weight = RNA / weight
 - weight = biopsy weight

Weight and RNA was summarised, RNA was devided by mean centered weight. Thus, we could compare total RNA results analyzed on different nanoDrop plates.

rRNA normalization: 
 1. Lambda expression was first calculated and stored in the data.frame "nf"
 2. Calculation of target expression
 3. Joined the lambda and target expression data.frames
 4. nf.w = expr / weight
 5. nf.w = nf.w / max(nf.w)
 6. nf.expr = log(expr / nf.w), nf.w = scale(nf.w) ?
 - expr = target expression 
 - weight = biopsy weight

To accurately measure changes in expression of rRNA, we added a fixed amount of Lambda to each total RNA sample during RNA extraction. Lambda expression and target (47S, 18S, 28S, 5.8S, 5S) were calculated in separate data frames, and then joined for normalization. In each data frame, we first created a normalization factor for weight (nf.w), a result of expression divided by weight. To make the nf.w comparable across the qPCR plates, we put it on a scale through the code nf.w = nf.w /max/(nf.w). This code divides the mean nf.w by the highest observed value in the variable nf.w.

Protein normalization:
 1. tpl = tp - bg
 2. tpl = tpl/max(tpl)
 3. signal = signal / max(signal)
 4. norm.sign = signal / tpl
 5. pool = mean(pool)
 6. norm.sign = norm.sign / pool
 - tpl = total protein per lane
 - tp = total protein (mean gray pixels in lanes)
 - bg = background (mean gray pixels between lanes)
 - signal = pixels per band

To create the tpl normalization factor, we measured the mean gray both within (tp) and between (bg) the lanes of each total protein stain blot with the ImageJ software. So, each lane and each space between lanes got a "mean gray" value. Subtracting background from total protein (tpl = tp-bg), and then using the resulting value divided by the maximum observed total protein in a lane per gel (tpl = tpl/max(tpl)), gave us a value normalized between gels. In addition, to account for variations in loading and unequal size in muscle tissue samples, we also normalized antibody signal by pools. The pools were mixes of total protein sample, 20ul from each sample, loaded into three lanes per gel. So, after normalizing signal to tpl (norm.sign = signal / tpl), we summarized the average signal from the pools (pool = mean(pool)), and divided the tpl normalized signal by average pool signal (norm.sign = norm.sign / pool).